---
title: "SIOP Case Study 2: Validation Report"
author: "Antigravity Coding Assistant"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    theme: united
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(dplyr)
library(readr)
library(ggplot2)
library(tidyr)
```

## 1. Executive Summary

This R Markdown document serves as the data validation report for **SIOP Case Study 2**. The objective of this analysis is to transform monthly SIOP forecast snapshots to compare the latest snapshot alongside prior snapshots (Resultant = 2023 P12, Lag 0 = 2023 P11, Lag 1 = 2023 P10) and merge actuals by `Item`, `Country`, and `Period` for forecast accuracy and bias reporting.

```{r load_data}
# Load the transformed data
csv_path <- "SIOP_Case_Study_2_Transformed_Data.csv"
if (!file.exists(csv_path)) {
  # Fallback to absolute path if needed
  csv_path <- "/Users/bhavya/Documents/Codex/2026-06-12/files-mentioned-by-the-user-case/outputs/siop_case_study_2/SIOP_Case_Study_2_Transformed_Data.csv"
}
df <- read_csv(csv_path)
```

The dataset contains **`r format(nrow(df), big.mark=",")`** rows and **`r ncol(df)`** columns.

---

## 2. Dataset Key Statistics

Here is a summary of the dimensions and metric totals from the transformed dataset:

```{r summary_stats}
summary_stats <- tibble(
  Metric = c(
    "Total Rows",
    "Distinct Items",
    "Countries Covered",
    "Periods Covered",
    "Total Actual Revenue",
    "Total Resultant Revenue Forecast (2023 P12)",
    "Total Lag 0 Revenue Forecast (2023 P11)",
    "Total Lag 1 Revenue Forecast (2023 P10)",
    "Total Actual Non-Revenue",
    "Total Actual Capex"
  ),
  Value = c(
    format(nrow(df), big.mark=","),
    format(n_distinct(df$Item), big.mark=","),
    as.character(n_distinct(df$Country)),
    as.character(n_distinct(df$Period)),
    format(sum(df$Actual_Revenue, na.rm=TRUE), big.mark=",", nsmall=2),
    format(sum(df$Resultant_Revenue_Forecast, na.rm=TRUE), big.mark=",", nsmall=2),
    format(sum(df$Lag0_Revenue_Forecast, na.rm=TRUE), big.mark=",", nsmall=2),
    format(sum(df$Lag1_Revenue_Forecast, na.rm=TRUE), big.mark=",", nsmall=2),
    format(sum(df$Actual_NonRevenue, na.rm=TRUE), big.mark=","),
    format(sum(df$Actual_Capex, na.rm=TRUE), big.mark=",")
  )
)

knitr::kable(summary_stats, caption = "Transformed Dataset Metrics Summary")
```

---

## 3. Revenue KPI Analysis by Country

Using the standard SIOP metrics:
- **Forecast Accuracy %** = $1 - \frac{|\text{Forecast} - \text{Actual}|}{\text{Actual}}$
- **Forecast Bias %** = $\frac{\text{Forecast} - \text{Actual}}{\text{Actual}}$

We aggregate the records by Country and calculate the KPIs for the Resultant, Lag 0, and Lag 1 snapshots:

```{r country_kpis}
country_summary <- df %>%
  group_by(Country) %>%
  summarize(
    Actual_Revenue = sum(Actual_Revenue, na.rm=TRUE),
    Resultant_Forecast = sum(Resultant_Revenue_Forecast, na.rm=TRUE),
    Lag0_Forecast = sum(Lag0_Revenue_Forecast, na.rm=TRUE),
    Lag1_Forecast = sum(Lag1_Revenue_Forecast, na.rm=TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    `Resultant Accuracy %` = round((1 - abs(Resultant_Forecast - Actual_Revenue) / Actual_Revenue) * 100, 2),
    `Resultant Bias %` = round(((Resultant_Forecast - Actual_Revenue) / Actual_Revenue) * 100, 2),
    `Lag 0 Accuracy %` = round((1 - abs(Lag0_Forecast - Actual_Revenue) / Actual_Revenue) * 100, 2),
    `Lag 0 Bias %` = round(((Lag0_Forecast - Actual_Revenue) / Actual_Revenue) * 100, 2),
    `Lag 1 Accuracy %` = round((1 - abs(Lag1_Forecast - Actual_Revenue) / Actual_Revenue) * 100, 2),
    `Lag 1 Bias %` = round(((Lag1_Forecast - Actual_Revenue) / Actual_Revenue) * 100, 2)
  )

knitr::kable(country_summary, caption = "Revenue Performance by Country and Lag")
```

### Visualizing Actual vs Forecast Revenue by Country
```{r chart_revenue, fig.width=10, fig.height=6}
country_long <- country_summary %>%
  select(Country, Actual = Actual_Revenue, Resultant = Resultant_Forecast, `Lag 0` = Lag0_Forecast, `Lag 1` = Lag1_Forecast) %>%
  pivot_longer(cols = -Country, names_to = "Metric", values_to = "Revenue")

ggplot(country_long, aes(x = Country, y = Revenue, fill = Metric)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Actual Revenue vs Forecast Snapshots by Country",
    y = "Revenue Quantity/Value",
    x = "Country",
    fill = "Version"
  )
```

---

## 4. Power BI Modeling Guide

To replicate this logic in Power BI:
1. **Fact Table**: Import `SIOP_Case_Study_2_Transformed_Data.csv`.
2. **Lag Selection**: Create a disconnected table named `Lag Selection` containing the column `Lag` with values: `"Resultant"`, `"Lag 0"`, and `"Lag 1"`.
3. **DAX Measures**:
   - **Lag Revenue Forecast**:
     ```dax
     Lag Revenue Forecast = 
     SWITCH(
         SELECTEDVALUE('Lag Selection'[Lag]), 
         "Resultant", SUM(SIOP[Resultant_Revenue_Forecast]), 
         "Lag 0", SUM(SIOP[Lag0_Revenue_Forecast]), 
         "Lag 1", SUM(SIOP[Lag1_Revenue_Forecast])
     )
     ```
   - **Revenue Forecast Accuracy %**:
     ```dax
     Revenue Forecast Accuracy % = 
     1 - DIVIDE(
         ABS([Lag Revenue Forecast] - SUM(SIOP[Actual_Revenue])), 
         SUM(SIOP[Actual_Revenue])
     )
     ```
   - **Revenue Forecast Bias %**:
     ```dax
     Revenue Forecast Bias % = 
     DIVIDE(
         [Lag Revenue Forecast] - SUM(SIOP[Actual_Revenue]), 
         SUM(SIOP[Actual_Revenue])
     )
     ```

---

## 5. Findings & Recommendations
- **Under-Forecasting**: Countries A, B, and C show massive negative bias (ranging from -72.9% to -94.9%), indicating a significant gap where forecasts under-represent actual demand.
- **Over-Forecasting**: Country D shows extreme positive bias (forecasts are 3.7x to 4.5x the actual demand), resulting in negative accuracy.
- **Action**: Implement machine learning or historical baseline corrections in the SIOP forecasting engine to correct systemic biases before submitting snapshots to BI.
